## Logistic regression (week 3)

Dataset is read from a local folder
```{r}
getwd()
setwd("/Users/jonnelintunen/Desktop/IODS-project/data")
alc <- read.table("alc.txt", header = T)
head(alc)
tail(alc)
```

### Selecting variables and hypothesis

Selected variables are studytime, failures, romantic and absences. Hypothesis are:
* high alcohol users use less time for studying
* high alc users have more failures
* high alc users are more usually men  
* high alc users have more absences 

### Logistic regression

```{r}
logistic_model <- glm(high_use ~ studytime + failures + sex + absences, data = alc, family = "binomial")
summary(logistic_model)
```
studytime, sex and absences are statistically significant (z-value over 2 which corresponds approx to P-value < 0.05). The bigger the z-value is the better:  z-value is Coefficient Estimate divided by Standard Error, so if z-value is big it means that Standard Error is small. 

```{r}
coef(logistic_model)
```

### Odds ratio
```{r}
library(dplyr)
OR <- coef(logistic_model) %>% exp
CI <- confint(logistic_model) %>% exp
cbind(OR, CI)
```
And again, studytime, sex and absences are statistically significant because CI doesn't include one.
Results for significant variables: 
* When studytime increases by one unit, alcohol consumption is 0.71 (0.51-0.97) times lower
* Being male increases the odds of being high alc user by 2.21 times (1.36-3.65) compared to females
* When absences increase by one unit, alcohol consumption is 1.09 (1.05-1.14) times higher 

### Binary predictions 
```{r}
new_model <- glm(high_use ~ studytime + sex + absences, data = alc, family = "binomial")
prob <- predict(new_model, type = "response")
alc <- mutate(alc, probability = prob)
alc <- mutate(alc, prediction = prob > 0.5)

table(high_use = alc$high_use, prediction = alc$prediction)
```
Model predicts 10 + 88 = 98 (26 %) cases wrongly from the total of 382 cases. 


