## Analysis of longitudinal data (week 6)

### BPRS dataset
```{r}
library(ggplot2)
getwd()
setwd("/Users/jonnelintunen/Desktop/IODS-project/data")
BPRSlong <- read.table("BPRSlong.txt", header = T)
BPRSlong$treatment <- factor(BPRSlong$treatment)
BPRSlong$subject <- factor(BPRSlong$subject)
ggplot(BPRSlong, aes(x = week, y = bprs, group = treatment, linetype = treatment)) + 
  geom_line()

```
It is quite hard to interpret this plot but it seems that in both treatment groups there is first a decrease in bprs values but in the treatment group 1 bprs value increases again at the end (week 8).

### RATS dataset
```{r}
library(ggplot2)
getwd()
setwd("/Users/jonnelintunen/Desktop/IODS-project/data")
RATSlong <- read.table("RATSlong.txt", header = T)
RATSlong$Group <- factor(RATSlong$Group)
ggplot(RATSlong, aes(x = Time, y = Weight, group = ID, linetype = Group)) + 
  geom_line()

```
We can see that there is more increase in weight with the rats in the groups 2 and 3. In the group 1 the  increase is rather insignificant. 



